### 2018-07-05 paper/5-implementations.tex
Minor revision

\simon{New subsection here? Or at least put this para after the next one.}
Using the abstractions built thus far, we have shown how to combine schedulers
with rebuilders to reproduce existing build systems. To us, the most interesting
build system as yet unavailable would combine a suspending scheduler with
constructive traces~--~providing a cloud-capable build system that is minimal,
and supports both early cutoff and monadic dependencies. Using our framework it
is possible to define and test such a system, which we call \Cloud \Shake. All
we need to do is replace \hs{restarting2} with \hs{suspending} in the definition
of \Bazel, as shown in Fig.~\ref{fig-cloud-implementations}.

@@

% Some build systems allow to exploit \emph{build task determinism} for efficient
% fetching of build results from the cache: one can simply hash the values of all
% transitive inputs of a key (which can be done efficiently by storing all hashes
% in a Merkle tree) and download the matching value from the cache, without the
% need to build any intermediate values. This can bring significant performance
% benefits in practice, but sacrifices the early cutoff property: indeed, with
% this approach addition of a new comment to one of the source files will
% invalidate all dependent keys.

>>>>

@@

Using the abstractions built thus far, we have shown how to combine schedulers
with rebuilders to reproduce existing build systems. To us, the most interesting
build system as yet unavailable would combine a suspending scheduler with
constructive traces~--~providing a cloud-capable build system that is minimal,
and supports both early cutoff and monadic dependencies. Using our framework it
is possible to define and test such a system, which we call \Cloud \Shake. All
we need to do is replace \hs{restarting2} with \hs{suspending} in the definition
of \Bazel, as shown in Fig.~\ref{fig-cloud-implementations}.

### 2018-07-05 paper/6-engineering.tex
Final tweaks
Minor revision

% Q: {Isn't ``early aborting'' automatically handled by short-circuiting
% withing monads, i.e. \hs{Nothing} or \hs{Left e} cancel out all subsequent
% computation?}
% NM says: The build system could chose an early abort monad, but the tasks are
% free in the monad, so you can't get the encoding to cross between them.

@@

\item[Volatility] Some tasks are defined to change in every execution, for
example \Excel provides a~``function'' \cmd{RANDBETWEEN} which produces a random
number in a specified range~--~on each recalculation it is defined to change.
Similarly, build systems like \Make and \Shake provide \emph{phony rules} which
are also volatile.
\end{description}
\simon{Is there a difference between non-determinism and volatility?  Perhaps
  one is accidental non-det while volatility is by-design non-det? But does the
  distinction matter?   Maybe for accidental non-det you are free to use an
  earlier cached result, whereas for volatity you should always recompute?
  Volatility is best handled by a dependency on RealWorld, adn we coudl say that right here.}

@@

Note that \hs{Task}~\hs{MonadPlus} is powerful enough to model dependency-level
non-determinism, for example, \cmd{INDIRECT("A" \& RANDBETWEEN(1,2))}, whereas
most build tasks in real-life build systems only experience a value-level
non-determinism. \Excel handles this example simply by marking the cell
volatile~--~an approach that can be readily adopted by any of our
implementations by introducing a special key \cmd{RealWorld} whose value is
changed between every run.

@@

\item[Frankenbuilds] A build is considered a
\emph{frankenbuild}~\cite{esfahani2016cloudbuild} if a value is calculated
locally, but something that depends on that key is pulled from the cache, and
the value calculated locally does not match what was previously calculated and
stored in the cloud. \simon{Don't understand. The explanation in 4.2.4 is better -- and I think describes
the same thing?}
Our implementations avoid this issue by storing complete
traces, but if a cloud build system was to only reference input nodes this
situation can arise.

@@

\noindent
In \Excel the user can alter the value produced by \cmd{B1} by either editing
the inputs of \cmd{A1} or \cmd{A2}, \emph{or} editing the formula in
\cmd{B1}~--~e.g. to \cmd{A1 - A2}. This pattern can be captured by describing
the rule producing \cmd{B1} as also depending on the value \cmd{B1-formula}.
The implementation can be given very directly in a \hs{Tasks}~\hs{Monad} as:

@@

\noindent
Namely, first look up the formula, then interpret it. It is not possible to
change dependencies based on the formula in a \hs{Tasks}~\hs{Applicative}, as
would be required, so instead the formula can be captured as a dependency
(but its value not used) and \emph{also} baked directly into the \hs{sprsh5}
function. \simon{Don't understand.}

The build systems that have precise self-tracking are all ones which use a
\emph{non-embedded domain specific language} to describe computations.
\simon{I think you mean that you can find dependencies with executing the function,
  which seems quite difference from ``backing it into'' the Haskell code.}
Those which make use of a full programming language, e.g. \Shake, are faced with
the challenge of implementing equality on arbitrary task functions. For such
build systems the incredibly pessimistic assumption of saying that any change to
the build system potentially changes any build rule can often be used~--~the
classic example being a makefile depending on itself.

@@

\noindent
In such cases \Excel will normally not execute anything, but if the
``Iterative Calculations'' feature is enabled will execute the formula for a
specified maximum number $N$ of times per calculation (where $N$ is a
setting that defaults to 100).

>>>>

@@

\item[Volatility] Some tasks are defined to change in every execution, for
example \Excel provides a~``function'' \cmd{RANDBETWEEN} which produces a random
number in a specified range~--~on each recalculation it is defined to change.
Similarly, build systems like \Make and \Shake provide \emph{phony rules} which
are also volatile. The main difference from non-deterministic tasks, is that
volatile tasks cannot be cached. They are best handled by introducing a special
key \cmd{RealWorld} whose value is changed between every run.
\end{description}

@@

Note that \hs{Task}~\hs{MonadPlus} is powerful enough to model dependency-level
non-determinism, for example, \cmd{INDIRECT("A" \& RANDBETWEEN(1,2))}, whereas
most build tasks in real-life build systems only experience a value-level
non-determinism. \Excel handles this example simply by marking the cell
volatile~--~an approach that can be readily adopted by any of our
implementations.

@@

% \item[Frankenbuilds] A build is considered a
% \emph{frankenbuild}~\cite{esfahani2016cloudbuild} if a value is calculated
% locally, but something that depends on that key is pulled from the cache, and
% the value calculated locally does not match what was previously calculated and
% stored in the cloud. \simon{Don't understand. The explanation in 4.2.4 is better -- and I think describes
% the same thing?}
% Our implementations avoid this issue by storing complete
% traces, but if a cloud build system was to only reference input nodes this
% situation can arise.

@@

\noindent
In \Excel the user can alter the value produced by \cmd{B1} by either editing
the inputs of \cmd{A1} or \cmd{A2}, \emph{or} editing the formula in
\cmd{B1}~--~e.g. to \cmd{A1 * A2}. This pattern can be captured by describing
the rule producing \cmd{B1} as also depending on the value \cmd{B1-formula}.
The implementation can be given very directly in a
\hs{Tasks}~\hs{Monad}~--~concretely, first look up the formula, then interpret
it:

@@

% It is not possible to
% change dependencies based on the formula in a \hs{Tasks}~\hs{Applicative}, as
% would be required, so instead the formula can be captured as a dependency
% (but its value not used) and \emph{also} baked directly into the \hs{sprsh5}
% function. \simon{Don't understand.}
% \simon{I think you mean that you can find dependencies with executing the function,
%   which seems quite difference from ``backing it into'' the Haskell code.}

\noindent
The build systems that have precise self-tracking are all ones which use a
\emph{non-embedded domain specific language} to describe computations.
Those which make use of a full programming language, e.g. \Shake, are faced with
the challenge of implementing equality on arbitrary task functions. For such
build systems the incredibly pessimistic assumption of saying that any change to
the build system potentially changes any build rule can often be used~--~the
classic example being a makefile depending on itself.

@@

In such cases \Excel will normally not execute anything, but if the
``Iterative Calculations'' feature is enabled will execute the formula for a
specified maximum number $N$ of times per calculation (where $N$ is a
setting that defaults to 100).
